---
title: "w9 and hbk bar chart dipterans by site"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
opts_knit$set(root.dir = "C:/Users/tyler/OneDrive/Documents/Huang Summer/Data Files/")

```

# Step 1: 
Set working directory and load files. I started with two separate data files, one with all of the data from HBK and one from all of the data from watershed 9 

```{r cars}
# start with packages needed
library("magrittr")
library("dplyr")
library("tidyr")
library("lubridate")
library("ggplot2")
```

# Step 2: 
Load data from watershed 9 into the data set, clean by changing the taxon and size as a variable, not an observation. 

```{r } 
ws9 <- read.csv("Sticky trap data - Watershed 9.csv")


names(ws9)
ws9_tidy <- ws9 %>% gather("Dipteran_large","Terrestrial_large","Caddisfly_large","Mayfly_large","Stonefly_large","Other_large","Dipteran_small","Terrestrial_small","Other_small", key ="invert_size", value = "count") %>% 
  separate(invert_size, into=c("invert", "size"), sep = "_") %>%  #making invertebrate type and size into separate variables
  
  
  #mutate changes things in data frames
  mutate(count_new = ifelse(is.na(count) == TRUE, 0, count)) %>% 
  dplyr::select(-count) %>%  #specifies which package to look for select in
  rename("trap" = "Trap..", "date" = "Date") %>% 
  mutate(date_lub = lubridate::mdy(date)) %>% 
  dplyr::select(-date) %>% 
  dplyr::select(date_lub, trap, invert, size, count_new) %>% 
  rename("date" = "date_lub") %>% 
 filter(date >= as.Date('2018-07-08')) #this line adjusts the dates so they'll be in the same range

```

Next, I put a label on all of the data so when I compare it to the data from the hubbard brook watershed, it'll be easy to distinguish. 
```{r }
 View(ws9_tidy)

ws9_tidy$site = "WS9"

ws9_sum <- ws9_tidy %>% group_by(date, invert, size, site) %>% summarize(count_sum = sum(count_new))  
```

# Step 3: 
Repeat the process of data cleaning with the data from HBK
```{r }
hbk <- read.csv("Sticky trap data - HBK.csv")

hbk_tidy <- hbk %>% gather("Dipteran_Large", "Terrestrial_Large" ,"Caddisfly_Large" ,
                           "Mayfly_Large","Stonefly_Large","Other_Large","Dipteran_Small" ,"Terrestrial_Small",
                           "Other_Small", key = "invert_size", value = "count_NA") %>% 
  
  #step 5, we're going to make invert and size into separate columns by piping this into the separate functio. 
  separate(invert_size, into = c("invert", "size"), sep = "_") %>% 
  
  #step 6, we're going to substitute the NA's in the count column with zeroes by piping into a for loop
  #and using the mutate function
  mutate(count = ifelse(is.na(count_NA)==TRUE, 0, count_NA)) %>% 

  
  
  #step 7, remove the count_NA column using select from dplyr.
  dplyr::select(-count_NA) %>% 
  
  #step 8, format the dates using dplyr
  mutate(date = lubridate::mdy(Date)) %>% 
  dplyr::select(-Date) %>% 
  dplyr::select(date, Trap, invert, size, count) 
```
 
 and the labelling to make it easy later. 
```{r }
 hbk_tidy$site = "HBK"

hbk_sum <- hbk_tidy %>% group_by(date, invert, size, site) %>% summarize(count_sum = sum(count))
head(hbk_sum)
```

# Step 4: Graphing
The first graph I made with this data counted the number of dipterans collected from HBK over time. Then, do the same for watershed 9
```{r}
#here's the graph for hbk's dipts alone
ggplot(hbk_sum %>% filter(invert == "Dipteran"), aes( date, count_sum)) + 
  geom_bar(stat = "identity")+ ggtitle("HBK")+ labs(x= "Date", y = "Number of Diptera")

#here's the graph we're going to use for ws9
ggplot(ws9_sum %>% filter(invert == "Dipteran"), aes(x = date, y = count_sum)) + 
  geom_bar(stat = "identity")+ ggtitle("WS9")+ labs(x= "Date", y = "Number of Diptera")

```

Then, we joined the two data sets together and graph them together. 
```{r}
#push the two dataframes together 

hbk_ws9 <- full_join(hbk_sum, ws9_sum)

#and here's how we put them on top of each other
ggplot(hbk_ws9 %>% filter(invert == "Dipteran"), aes(fill = site, x = date, y = count_sum)) +
  geom_bar(stat = "identity")

```


